\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip Spatiotemporal modeling of \emph{Mycobacterium tuberculosis} transmission risks using environmental, clinical, and patient movement data}

### 2023-08-10 doc/paper/supplements.tex

@@

Tracking data required preprocessing before it could be used for spatiotemporal modeling. The reason is that there were considerably more tracking IDs than what would be expected from the number of registered patients in the clinical database. Clinical attendees most likely contributed multiple tracks for several reasons: 
\begin{itemize}
    \item A person moved outside the range of the video sensors and then re-entered. 
    \item An individual may temporarily not have been recognized as a person by the tracking software, \eg because the person bowed.
    \item A person may have been temporarily hidden behind another person.
    \item Moving objects such as opening doors may have been recognized as a person, adding a few noisy tracks into the data.
\end{itemize}
We proceeded in two steps to link back together tracking IDs that most likely belonged to the same person. In the first step, we filtered noise (\ie IDs of less than 10s and a total moving distance of less than 2m) and automatically matched IDs that were less than 10s and 3m apart while moving (\ie IDs anywhere in the clinic) or less than 60s and 0.5m apart while sitting (\ie two IDs ending/starting in a seating area of the clinic). In the second step, we build on the automatically matched data and performed an additional manual linkage. The aim was to link IDs that were further apart but still most likely belonged together. Through manual inspection, we could better assess whether a link resulted in a movement that made sense and preclude better alternatives. A Shiny application and a matching manual/wiki (\url{https://github.com/nbanho/clinic-transmission/wiki/Matching-Manual}) was developed to assist the students that performed the manual linkage.

The Shiny app is shown in \Cref{fig:shiny-app}. In the right panel, a colored floor plan of the clinic is shown with the different areas delineated. Seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with an open circle indicating the start and a crossed circle indicating the end of a track. The currently selected track is shown in black, possible links are shown in colors (here only one possible link shown in red), and alternatives to the selected possible link are shown in light grey. Below the floor plan is a table showing descriptive information about the possible links in relation to the currently selected track: (1)~the patient ID (Pid) of the selected track and observation ID (Oid) of the possible link, (2)~the height difference between the end of the Pid and end of the Oid track, (3)~ the standing height (\ie maximum height of a track) difference, (4)~the duration of the possible link, (5)~the time difference, and (6)~the distance between the end of the selected track and the start of the selected possible link. Possible links are sorted in ascending order by the time difference.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{doc/paper/shiny-tool-screenshot.png}
    \caption{Screenshot of the Shiny app used for manual linking IDs that probably belong together.}
    \label{fig:shiny-app}
\end{figure}

>>>>

\subsection{Tracking data}

\begin{figure}
    \centering
    \includegraphics{trac}
    \caption{Caption}
    \label{fig:enter-label}
\end{figure}

@@

Tracking data required preprocessing before it could be fully used for spatiotemporal modeling. The reason is that there were considerably more tracking IDs than what would be expected from the number of registered patients in the clinical database. Clinical attendees most likely contributed multiple tracks for several reasons: 
\begin{itemize}
    \item A person moved outside the range of the video sensors and then re-entered. 
    \item An individual may temporarily not have been recognized as a person by the tracking software, \eg because the person bowed.
    \item A person may have been temporarily hidden behind another person.
    \item Moving objects such as opening doors may have been recognized as a person, adding a few noisy tracks into the data.
\end{itemize}
We proceeded in two steps to link back together tracking IDs that most likely belonged to the same person. In the first step, we filtered noise (\ie IDs of less than 10s and a total moving distance of less than 2m) and automatically matched IDs that were less than 10s and 3m apart while moving (\ie IDs anywhere in the clinic) or less than 60s and 0.5m apart while sitting (\ie two IDs ending/starting in a seating area of the clinic). In the second step, we build on the automatically matched data and performed an additional manual linkage. The aim was to link IDs that were further apart but still most likely belonged together. Through manual inspection, we could better assess whether a link resulted in a movement that made sense and preclude better alternatives. A Shiny application and a matching manual/wiki (\url{https://github.com/nbanho/clinic-transmission/wiki/Matching-Manual}) was developed to assist the students that performed the manual linkage.

The Shiny app is shown in \Cref{fig:shiny-app}. In the right panel, a floor plan of the clinic is shown in (x,y)-space. Different areas of the clinic are delineated by colors, \eg seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with an open circle indicating the start and a crossed circle indicating the end of a track. The currently selected track is shown in black, possible links are shown in colors (here only one possible link shown in red), and alternatives to the selected possible link are shown in light grey. Below the floor plan is a table showing descriptive information about the possible links in relation to the currently selected track: (1)~the patient ID (Pid) of the selected track and observation ID (Oid) of the possible link, (2)~the height difference between the end of the Pid and end of the Oid track, (3)~ the standing height (\ie maximum height of a track) difference, (4)~the duration of the possible link, (5)~the time difference, and (6)~the distance between the end of the selected track and the start of the selected possible link. Possible links are sorted in ascending order by the time difference.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics[width=\linewidth]{doc/paper/shiny-tool-screenshot.png}
    \caption{Screenshot of the Shiny app used for manual linking IDs that probably belong together.}
    \label{fig:shiny-app}
\end{figure}
